% We develop a likelihood free inference procedure for conditioning a probabilistic model on a predicate.

% A predicate is a Boolean valued function which expresses a yes/no question about a domain.

% Our contribution, which we call predicate exchange, 
% constructs a softened predicate which takes value in the unit interval [0, 1] as opposed to a simply true or false. 

% Intuitively, 1 corresponds to true, and a high value (such as 0.999) corresponds to ``nearly true'' as determined by a distance metric.

% We define Boolean algebra for soft predicates,  such that they can be negated, conjoined and disjoined arbitrarily.

% A softened predicate can serve as a tractable proxy to a likelihood function for approximate posterior inference.

% However, to target exact inference, we temper the relaxation by a temperature parameter, and add a accept/reject phase use to replica exchange Markov Chain Mont Carlo, which exchanges states between a sequence of models conditioned on predicates at varying temperatures.

% We describe a lightweight implementation of predicate exchange that it provides a language independent layer that can be implemented on top of existingn modeling formalisms.

% Conditioning a model on a predicate incorporates information into the model, but often renders the likelihood function intractable.
% To sample from them likelihood-free inference is required, 

% ---
Conditioning models on predicates incorporates information
into models. 
Models conditioned on predicates rarely have a tractable likelihood; sampling from them requires likelihood-free inference.
However, traditional approaches to likelihood-free 
inference focus on predicates involving observed data.
We develop a likelihood free inference procedure for a broader class of
predicates called predicate exchange. 
Predicate exchange constructs \emph{soft} predicates, which return values in a continuous Boolean algebra: the unit interval with continuous logical connectives.
A soft predicate can serve as a proxy likelihood function in inference,
but introduces an approximation error.
The error depends on a temperature setting, such that zero-temperature corresponds to the original predicate, but higher temperatures mix faster.
To draw samples from the model conditional on the original predicate,
we construct Markov chains at different temperatures and use 
replica exchange to swap between chains.
We implement predicate exchange as a language-independent layer
that can be implemented on top of existing probabilistic programming formalisms. 
We demonstrate the approach on sequence models of health and probabilistic renderers.



% 